\Large \bf Cyclone: Warp Speed Replication for Persistent Key Value Stores

### 2017-07-27 paper/paper.tex
Additions.

The first key contribution of Cyclone is demonstrating the benefit of a small
amount of non-volatile memory when combined with the flash SSD based persistent
logs.  The primary cause of slowdown when turning on a write ahead log on a
single machine in Figure~\ref{fig:problem} is the need to do synchronous block
IO for every update. For small to moderately sized updates, this leads to severe
write amplification as a 4KB page must be synchronously written to the log file
on flash. Cyclone avoids this by making use of directly attached persistent
memory on the server. This can be in the form of NVDIMMs (DIMMs with an added
ultracapacitor to dump state to a small amount of attached flash) or newer forms
of persistent memory~\cite{pmfs}. For current deployments, NVDIMMs are expensive
- retailing at approximately 2X the price of a standard DIMM [find ref]. This
means that they will likely be available in small quantities. We compensate for
this by periodically draining the NVDIMM log into a log placed on a standard SSD
on the system.

@@

\section{Storage}
\label{sec:storage}
Storage devices such as flash-based SSDs export a block IO interface. Appending
a small update to the write-ahead log requires a synchronous update of an entire
4KB page. This leads to a high baseline latency at even low load, as shown in
Figure~\ref{fig:problem}.  On the other hand small updates (in the order of 100s
of bytes for both the key and value taken together) are commonplace in many
workloads [find refs]. This problem leads many system designers to
underprovision key-value store shards ensuring that it operates at moderate to
high load, using group-commit to batch additions to the log. They therefore pay
a price in latency to ensure the bandwidth to storage is fully utilized to
prevent it from becoming a bottleneck.

Cyclone adopts a different strategy by making use of novel memory technologies
to obviate the need to make this tradeoff in the first place. We assume a small
amount of non-volatile memory directly addressable by the CPU rather than placed
behind a block oriented IO interface. Log appends are done to the ``NVM log''
placed in this non-volatile memory. While we designed Cyclone to leverage new
non-volatile memory technologies that might be available in the
future~\cite{3dxpoint} current non-volatile memory options such are NVDIMMs
are more expensive that DRAM. Therefore, for Cyclone to be economical
we cannot assume that a sufficient amount of NVM is available to hold the entire
log. The log entries can only be garbage collected when the in-memory component
of the key value store is checkpointed to secondary storage. The growing size of
available DRAM means that also provisioning sufficient NVDIMMs to hold the
corresponding log of updates is not feasible given that NVDIMMs cost at least
twice as much as volatile DRAM. We therefore resort to using a second larger log
level placed on a flash SSD. Entries are drained from the NVM log to this flash
log in conveniently large units that are a multiple of 4KB - the optimum IO unit
for flash SSDs. This ensures we get the best
possible throughput from a flash device without paying the price of synchronous
IO for small key-value pairs. At the same time, the amount of non-volatile
memory required is small enough to not add undue cost of the server.

@@

Using non-volatile memory to cache writes to flash is also employed in the
internal architecture of flash SSDs - including the SSD we use in the evaluation
of the paper. This ``write cache'' however does not help with small appends to
the write ahead log, as experiments with Rocksdb's own write-ahead log has shown
in this paper.

>>>>

The first key contribution of Cyclone is demonstrating the benefit of a small
amount of non-volatile memory when combined with the flash SSD based persistent
logs.  The primary cause of slowdown when turning on a write ahead log on a
single machine in Figure~\ref{fig:problem} is the need to do synchronous block
IO for every update. Mordern NVMe SSDs advertise a synchronous write latency of
the order of around 20us, capping throughput at 50K ops/sec, even with write coalescing 
in the SSDs own internal write cache. Cyclone avoids this by making use of directly attached persistent
memory on the server. This can be in the form of NVDIMMs (DIMMs with an added
ultracapacitor to dump state to a small amount of attached flash) or newer forms
of persistent memory~\cite{pmfs}. For current deployments, NVDIMMs are expensive
- retailing at approximately 2X the price of a standard DIMM [find ref]. This
means that they will likely be available in small quantities. We compensate for
this by periodically draining the NVDIMM log into a log placed on a standard SSD
on the system.

@@

\section{Storage}
\label{sec:storage}
Storage devices such as flash-based SSDs export a block IO interface. Appending
a small update to the write-ahead log requires a synchronous access to the SSD. Even with 
write coalescing using a write cache on the SSD, the round  trip time to a current generation NVMe 
SSD such as the one we use in this paper~\cite{ssd_spec}, is of the order of 20 us, limiting the throughput to under 50K ops/sec.

This leads to a high baseline latency at even low load, as shown in
Figure~\ref{fig:problem}. This problem leads many system designers to
underprovision key-value store shards ensuring that it operates at moderate to
high load, using group-commit to batch additions to the log. They therefore pay
a price in latency to ensure the bandwidth to storage is fully utilized to
prevent it from becoming a bottleneck.

Cyclone adopts a different strategy by making use of novel memory technologies
to obviate the need to make this tradeoff in the first place. We assume a small
amount of non-volatile memory directly addressable by the CPU rather than being placed
behind a block oriented IO interface. Log appends are done to the ``NVM log''
placed in this non-volatile memory. While we designed Cyclone to leverage new
non-volatile memory technologies that might be available in the
future~\cite{3dxpoint} current non-volatile memory options such are NVDIMMs
are more expensive that DRAM. Therefore, for Cyclone to be economical
we cannot assume that a sufficient amount of NVM is available to hold the entire
log. The log entries can only be garbage collected when the in-memory component
of the key value store is checkpointed to secondary storage. The growing size of
available DRAM means that also provisioning sufficient NVDIMMs to hold the
corresponding log of updates is not feasible given that NVDIMMs cost at least
twice as much as volatile DRAM. We therefore resort to using a second larger log
level placed on a flash SSD. Entries are drained from the NVM log to this flash
log in conveniently large units that are a multiple of 4KB - the optimum IO unit
for flash SSDs. This ensures we get the best
possible throughput from a flash device without paying the price of synchronous
IO for small key-value pairs. At the same time, the amount of non-volatile
memory required is small enough to not add undue cost of the server.

@@
